---
title: "LAGOS Spatial Analysis"
author: "Matthew Ross"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(ggplot2)
library(ggthemes)
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(version='1.087.1',dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
lagos <- lagosne_load('1.087.1')

#Grab the lake centroid info
lake_centers <- lagos$locus

load('lake_centers.Rdata')
```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100)

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r }
#Your code here

#load all states
states <- us_states()

#filter out just Iowa and Illinois
I_states <- states %>%
  filter(name %in% c('Illinois','Iowa')) %>%
  st_transform(2163)

#Show interactive map
mapview(I_states)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
##combined? How does this compare to Minnesota?

```{r}
#Subset lakes in Iowa and Illinois
I_lakes <- spatial_lakes[I_states,]

summarise(I_lakes, n())
summarise(minnesota_lakes, n())

#There are 16,466 in Illinois and Iowa combined. In comparison, there are 29,038
#lakes in Minnesota alone.

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
#Another way of doing this... 
#subset iowa and minnesota
iowa_minnesota <- states %>%
  filter(name %in% c('Iowa', 'Minnesota')) %>%
  st_transform(2163)

#subset just names, need to filter out NAs which are lakes in other states
iowa_minnesota_lakes <- spatial_lakes %>%
  st_join(., iowa_minnesota) %>%
  filter(!is.na(name))

#look at the data, names should be iowa and minnesota
table(iowa_minnesota_lakes$name)

#plot histograms of the two states
ggplot(iowa_minnesota_lakes, aes(lake_area_ha)) +
  geom_histogram(binwidth = 0.1, color = 'black', fill = 'gray85') + 
  scale_x_log10() + 
  xlab('Lake Size (ha)') + 
  ylab('Count')+
  theme_few()+
  facet_wrap(~name, dir = 'v', scales = 'free_y')

# #Subset just Iowa (without Illinois)
# Iowa <- states %>%
#   filter(name == 'Iowa') %>%
#   st_transform(2163) 
# 
# #Subset lakes that are just in Iowa
# Iowa_lakes <- spatial_lakes[Iowa,] %>%
#   mutate(state = 'Iowa')
# 
# #Add state name to Minnesota_lakes
# minnesota_lakes <-spatial_lakes[minnesota,] %>%
#   mutate(state = 'Minnesota')
# 
# #rbind the two datasets together
# MinIowa_lakes <- rbind(Iowa_lakes, minnesota_lakes)
# 
# #Create histograms comparing lake size for Minnesota and Iowa
# ggplot(data = MinIowa_lakes, aes(lake_area_ha)) + 
#   geom_histogram(binwidth = 0.1, color = 'black', fill = 'gray85')+ 
#   xlab('Lake Size (ha)') + 
#   ylab('Count')+
#   scale_x_log10()+
#   theme_few()+
#   facet_wrap(~state, dir = 'v')



```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
#Making a map of all Illinois and Iowa lakes colored by area
#slicing out the first 1000 lakes to make code run faster
I_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')

```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 


