---
title: "Lake Water Quality Analysis"
author: "Matthew Ross"
date: "9/17/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(lubridate) #For dealing with date and time
library(scales) #for graphing
library(ggthemes) #for plotting
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
lagosne_get(dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
names(nutr)
```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')

mapview(spatial_200)

```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')
```


# Class work

## 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for
sites with at least 200 observations?

- Here, I just want a plot of chla vs secchi for all sites 

```{r}
#Your code here
ggplot(mean_values_200, aes(x = mean_secchi, y = mean_chl))+
  geom_point(colour = 'black', fill = 'chartreuse4', size = 2,
             shape = 21)+
  theme_few()+
  scale_y_log10()+
  annotation_logticks(base = 10, sides ='l') +
  theme(axis.ticks.length = unit(-0.25, 'cm'), 
        axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5),"cm")),
        axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm"))) +
  xlab('Mean Secchi Disk Depth') + 
  ylab('Mean Chlorophyll A') 
  
```


## Why might this be the case? 
#At deeper depths, chlorophyll A cannot grow? 


## 2) What states have the most data? 

### 2a) First you will need to make a lagos spatial dataset that has the total number of counts per site.

```{r}
## Your code here

#count all measurements per lake 
chla_count <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

#looking at the structure of chla_count
str(chla_count)

#Make chla_count a spatial dataset by joining to spatial_lakes
spatial_count <- inner_join(spatial_lakes,chla_count,
                          by='lagoslakeid')

```


### 2b) Second, you will need to join this point dataset to the us_boundaries data. 

```{r}
## Your code here

#Call all states
states <- us_states()

#Join states to spatial_count
states_count <- spatial_count %>%
  st_join(., states, by = 'lagoslakeid')
```


### 2c) Then you will want to group by state and sum all the observations in that
state and arrange that data from most to least toatl observations per state. 

```{r}
## Your code here. 
states_count %>%
  group_by(state_name) %>%
  summarize(total_obs = sum(count)) %>%
  arrange(desc(total_obs))

## The state with the most observations is Minnesota with 78,455 observations
```

##3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200 
observations?

```{r}
## Your code here
mapview(mean_spatial, zcol = 'mean_secchi')

#From the map, it appears that lakes in the northeast tend to have higher mean secchi depths than lakes in the midwest

```


